---
title: "Crime"
author: "Juan Nunez"
date: "April 1, 2019"
output:
  word_document: default
  html_document: default
---

Set working directory.

```{r}
setwd("C:\\Users\\Juan Nunez\\Desktop\\DATA_Capstone")
```

Open packages.

```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)
```


Opening the crime dataset.

```{r}
crime <- read_csv("DATASETS\\Crime.csv")
```

Now let's move to the next dataset.

```{r}
crime2 <- crime
```


Let's see the top of this dataset.

```{r}
head(crime2)
```

Let's add underscores.

```{r}
colnames(crime2) = gsub(" ", "_",colnames(crime2))
```

Let's see

```{r}
head(crime2)
```

Let's change Dispatch_Date/Time name.

```{r}
colnames(crime2)[colnames(crime2)=="Dispatch_Date/Time"] <- "Date_and_Time"
```

Let's see.

```{r}
head(crime2)
```

Saving a new dataset.

```{r}
crime3 <- crime2
```

```{r}
head(crime3)
```


Now we create a separate matrix with the dates and times.

```{r}
library(data.table)

```

We split the date and time variable into tree parts whenever there is a space.

```{r}
setDT(crime3)[, paste0("Date_and_Time", 1:3) := tstrsplit(Date_and_Time, " ")]
```

```{r}
class(crime3)
```

Turn crime3 into a data frame.

```{r}
crime3 <- as.data.frame(crime3)
```


```{r}
head(crime3)
```

Below I add the PM to time variable.

```{r}
crime3$Time = paste(crime3$Date_and_Time2, crime3$Date_and_Time3, sep=" ")
```

Now we have a variable for the time and another for the date.

```{r}
head(crime3)
```

I create a date variable.

```{r}
class(crime3$Date_and_Time1)
```

```{r}
colnames(crime3)[colnames(crime3)=="Date_and_Time1"] <- "Date"
```


```{r}
head(crime3)
```

Convert to date format.

```{r}
crime3$Date <- as.Date(crime3$Date, "%m/%d/%Y")
```

```{r}
class(crime3$Date)
```

I add the dataset constant to identify this data in a merge.

```{r}
crime3['DATNUM']=2
```


```{r}
head(crime3)
```

Below I fix the district numbers variable.

```{r}
table(crime3$Police_District_Number)
```

```{r}
crime3$SubAgency_n <- crime3$Police_District_Number
```

```{r}
crime3$SubAgency_n[crime3$SubAgency_n== "TPPD"] <- "OTHER"
crime3$SubAgency_n[crime3$SubAgency_n== "OTHER"] <- "Other"
table(crime3$SubAgency_n)
```

```{r}
head(crime3)
```

```{r}
table(crime3$Crime_Name1)
table(crime3$Crime_Name2)
```

Below I add a constant for violation type to the whole dataset.

```{r}
crime3['V_type']=2
```

Those categories that are considered violent will have a 3.

```{r}
crime3$V_type[crime3$Crime_Name2== "Aggravated Assault" | crime3$Crime_Name2== "Fondling" | crime3$Crime_Name2== "Forcible Rape" | crime3$Crime_Name2== "Intimidation" | crime3$Crime_Name2== "Kidnapping/Abduction" | crime3$Crime_Name2== "Weapon Law Violations" | crime3$Crime_Name2== "Simple Assault" | crime3$Crime_Name2== "Sexual Assault With An Object"  | crime3$Crime_Name2== "Robbery"  | crime3$Crime_Name2== "Murder and Nonnegligent Manslaughter" | crime3$Crime_Name2== "Forcible Sodomy" | crime3$Crime_Name2== "Forcible Fondling"] <- 3



```

Those categories that are considered drug-related will have a 4.

```{r}
crime3$V_type[crime3$Crime_Name2== "Drug Equipment Violations" | crime3$Crime_Name2== "Drug/Narcotic Violations" | crime3$Crime_Name2== "Drunkenness"  ] <- 4
```

Below are the categories involved.

```{r}
table(crime3$V_type)
```

crime4 will be the merge dataset for the map.  It is a subset of crime3

```{r}
crime4 <- crime3
```

```{r}
head(crime4)
```



```{r}
crime4 <- crime4 %>% select(Date, DATNUM, Latitude, Longitude, SubAgency_n, V_type)
```

```{r}
table(crime4$V_type)
```

I save both the full and the subset data to csv to use in Tableau.

```{r}
write.csv(crime4, file = "crime4.csv",row.names=FALSE, na="")
```

```{r}
write.csv(crime3, file = "crime3.csv",row.names=FALSE, na="")
```


```{r}
class(crime3$Time)
```

```{r}
crime3$Time <- strptime(crime3$Time, "%I:%M:%S %p")
```

```{r}
head(crime3)
```

```{r}
crime3$Time_n <- format(crime3$Time,"%H:%M:%S")
```

```{r}
head(crime3)
```

```{r}
crime3$V_type_n[crime3$V_type == 2] <- "Other"
crime3$V_type_n[crime3$V_type == 3] <- "Violent"
crime3$V_type_n[crime3$V_type == 4] <- "Drugs"

```

```{r}
head(crime3)
```


```{r}
crime3$V_type_n <- factor(crime3$V_type_n, levels = c("Drugs", "Violent", "Other"))
table(crime3$V_type_n)
table(crime3$V_type)
```

```{r}
library(ggthemes)
```

```{r}
class(crime3$Time_n)
```

```{r}
head(crime3)
```


```{r}
class(crime3$Time_n)
```



```{r}
library(hms)
library(lubridate)
crime3$Time_n <- as.hms(crime3$Time_n)
```

```{r}
class(crime3$Time_n)
```


```{r}
table(crime3$Victims)
```

```{r}
class(crime3$Victims)
```

```{r}
crime3$Victims_n[crime3$Victims == 1] <- "1"
crime3$Victims_n[crime3$Victims == 2] <- "2"
crime3$Victims_n[crime3$Victims > 2] <- ">2"
table(crime3$Victims_n)
```

```{r}
crime3$Victims_n <- factor(crime3$Victims_n, levels = c("1", "2", ">2"))
table(crime3$Victims_n)
table(crime3$Victims)
```


```{r}
table(crime3$Crime_Name1)
```

```{r}
crime3$name <- crime3$Crime_Name1
crime3$name[crime3$name == "Not a Crime"] <- "Other"
table(crime3$name)
```

```{r}
table(crime3$Place)
```

```{r}
table(crime3$Zip_Code)
```

```{r}
write.csv(crime3, file = "crime3.csv",row.names=FALSE, na="")
```

```{r}
crime_n <- crime3
```

```{r}
crime3 <- subset(crime3, select=-c(Time))
```



```{r}
#filter(!is.na(Gender_n)) %>%
crime3 %>%  filter(SubAgency_n != "Other") %>% 
ggplot() +
aes(x = SubAgency_n) +
geom_bar() +
labs(fill = "Victims", y = "# of arrests", x ="District", title = "Number of Arrests per District") +
theme_stata()
```

```{r}
class(crime3$Victims)
```

```{r}
table(crime3$name)
```


```{r}
crime3 %>%  filter(SubAgency_n != "Other") %>%  filter(!is.na(name)) %>% 
ggplot() +
aes(x = SubAgency_n , fill = name) +
geom_bar() +
labs( y= "Arrests (#)" , fill = "Crime Type", x ="District", title = "Number of Crimes per District by Type") +
scale_fill_discrete(labels = c("Person", "Property", "Society", "Other")) +
theme_stata()
```


```{r}
north <- 39.36
south <- 38.92
west <- -77.55
east <- -76.88
```

```{r}

crime3$V_type_s <- crime3$V_type
crime3$V_type_s[crime3$V_type_s == 2] <- NA

table(crime3$V_type)
table(crime3$V_type_s)
```


```{r}
crimeplot <- crime3 %>% filter(Longitude>west,Longitude<east,Latitude<north,Latitude>south) %>% filter(!is.na(V_type_s)) %>%
ggplot() + 
geom_point(aes(x = Longitude, y = Latitude, col = as.factor(V_type_s)),alpha = .2, size = .9)+
labs(title = "Violence and Drugs Map", col = "Crime Type") 


print(crimeplot  +scale_colour_manual(labels = c("Violence", "Drugs"), values = c( "Red", "Green")))
```


```{r}
table(crime3$V_type_n)
```


```{r}
table(crime3$Victims_n)
```


```{r}
table(crime3$SubAgency_n)
```



```{r}
table(crime3$name)
```







```{r}
crime3 %>% filter()
```


