---
title: "MC_ENROLL"
author: "Juan Nunez"
date: "April 4, 2019"
output:
  word_document: default
  html_document: default
---


I set the working directory.

```{r}
setwd("C:\\Users\\Juan Nunez\\Desktop\\DATA_Capstone")
```

I open the packages.

```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)
```

I open the dataset.

```{r}
mc_enroll <- read_csv("DATASETS\\Montgomery_College_Enrollment_Data.csv")
```

Now let's look at the top dataset


```{r}
dim(mc_enroll)
head(mc_enroll)
```


Below is a code to see the entire dataset in terms of categories within variables.

```{r}
lapply(mc_enroll,table)
```


Below I add an underscore to the names of the columns.


```{r}
mc_enroll2 <- mc_enroll
colnames(mc_enroll2) = gsub(" ", "_",colnames(mc_enroll2))
```


I create a dummy for whether they are continuing or new to MC

```{r}
mc_enroll2['MC_TRUE']= "Yes"
mc_enroll2$MC_TRUE[mc_enroll2$Student_Type != "Continuing"] <- "No"
table(mc_enroll2$MC_TRUE)
```


Take out the unknowns from Race Ethnicity and Gender

```{r}
mc_enroll2$HISPANIC <- mc_enroll2$Ethnicity
mc_enroll2$HISPANIC[mc_enroll2$Ethnicity == "Unknown"] <- NA
table(mc_enroll2$HISPANIC)
```



```{r}
mc_enroll2$RACE <- mc_enroll2$Race
mc_enroll2$RACE[mc_enroll2$Race == "Unknown"] <- NA
table(mc_enroll2$RACE)
```



```{r}
mc_enroll2['SEX']= "F"
mc_enroll2$SEX[mc_enroll2$Gender == "Male"] <- "M"
mc_enroll2$SEX[mc_enroll2$Gender == "Unknown"] <- NA
table(mc_enroll2$SEX)
```



Below I create a 3 category variable for MC campus.

```{r}

mc_enroll2['MC_CAMPUS']= "Other"
mc_enroll2$MC_CAMPUS[mc_enroll2$`Attending_Takoma_Park/SS` == "Yes"] <- "TP - SS"
mc_enroll2$MC_CAMPUS[mc_enroll2$Attending_Rockville  == "Yes"] <- "RV"
mc_enroll2$MC_CAMPUS[mc_enroll2$Attending_Germantown  == "Yes"] <- "GT"
table(mc_enroll2$MC_CAMPUS)


```


Below I make unknown NA in the AGE variable

```{r}
mc_enroll2$AGE_GROUP_fixed <- mc_enroll2$Age_Group
mc_enroll2$AGE_GROUP_fixed[mc_enroll2$AGE_GROUP_fixed == "Unknown"] <- NA
table(mc_enroll2$AGE_GROUP_fixed)
```



Below I create a dummy for MCPS.

```{r}
mc_enroll2['MCPS_DUM']= "Other"
mc_enroll2$MCPS_DUM[mc_enroll2$HS_Category == "MCPS"] <- "MCPS"
table(mc_enroll2$MCPS_DUM)
```


Below I create a dummy for State

```{r}
mc_enroll2['STATE_DUM']= "Other"
mc_enroll2$STATE_DUM[mc_enroll2$State == "MD"] <- "MD"
table(mc_enroll2$STATE_DUM)
```


I fix the zip codes.

```{r}
class(mc_enroll2$ZIP)
mc_enroll2$ZIP <- as.character(mc_enroll2$ZIP)
class(mc_enroll2$ZIP)

```

```{r}
table(mc_enroll2$ZIP)
```

```{r}
mc_enroll2$ZIP[mc_enroll2$ZIP == "926"] <- "00926"
mc_enroll2$ZIP[mc_enroll2$ZIP == "2050"] <- "02050"
mc_enroll2$ZIP[mc_enroll2$ZIP == "6830"] <- "06830"
mc_enroll2$ZIP[mc_enroll2$ZIP == "7102"] <- "07102"
table(mc_enroll2$ZIP)

```


Clean County in MD.


```{r}
mc_enroll2['MD_COUNTY']= "Other"
mc_enroll2$MD_COUNTY[mc_enroll2$County_in_MD == "Montgomery"] <- "Montgomery"
head(mc_enroll2)
table(mc_enroll2$MD_COUNTY)
```


Below I save teh dataset to csv.

```{r}
write.csv(mc_enroll2, file = "mc_enroll2.csv",row.names=FALSE, na="")
```

```{r}
table(mc_enroll2$Student_Type)
```




```{r}
table(mc_enroll2$Student_Status)
```


```{r}
table(mc_enroll2$RACE)
```

```{r}
table(mc_enroll2$Race)
```

```{r}

mc_enroll2$RACE <- mc_enroll2$Race

mc_enroll2$RACE[mc_enroll2$RACE == "Multi-Race"] <- "Other"
mc_enroll2$RACE[mc_enroll2$RACE == "Native American"] <- "Other"
mc_enroll2$RACE[mc_enroll2$RACE == "Pacific Islander"] <- "Other"
mc_enroll2$RACE[mc_enroll2$RACE == "Unknown"] <- "Other"

table(mc_enroll2$RACE)

```

```{r}
table(mc_enroll2$AGE_GROUP_fixed)
```

```{r}
mc_enroll2$RACE_n <- factor(mc_enroll2$RACE, levels = c("White", "Black", "Hispanic", "Asian", "Other"))
table(mc_enroll2$RACE_n)
table(mc_enroll2$RACE)
```

```{r}
table(mc_enroll2$MC_CAMPUS)
```

```{r}
mc_enroll2 %>% filter(!is.na(AGE_GROUP_fixed)) %>%
ggplot(aes(x = AGE_GROUP_fixed, fill = MC_CAMPUS)) +
geom_bar() +
labs(title = "Enrollment by Age", x= "Age Group", y = "Enrollment") +
guides(fill=guide_legend(title = "Campus")) +
scale_fill_discrete(labels = c("Germantown", "Rockville", "Takoma Park"))

```

```{r}

mc_enroll2 %>%  filter(!is.na(MC_CAMPUS)) %>%
ggplot() +
aes(x = MC_CAMPUS , fill = Student_Status) +
geom_bar(position = 'dodge') +
scale_fill_discrete(labels = c("Full-Time", "Part-Time")) +
labs(fill = "Status", y = "Enrollment", x ="Campus", title = "Enrollment by Campus and Status") +
scale_x_discrete(labels = c("Germantown" , "Rockville", "Takoma Park"))
```

```{r}

```











