---
title: "Capstone_analyses"
author: "Juan Nunez"
date: "April 1, 2019"
output:
  word_document: default
  html_document: default
---

Set working directory.

```{r}
setwd("C:\\Users\\Juan Nunez\\Desktop\\DATA_Capstone")
```

Open packages

```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)
```

Open data:

```{r}
#traffic <- read_csv("DATASETS\\Traffic_Violations.csv")
##save(traffic, file = "traffic.Rda")
load(file = "traffic.rda")
```

Look at the top of the data.

```{r}
head(traffic)

```


Dimensions of the traffic dataset.

```{r}
dim(traffic)

```


Below I make a copy of traffic.

```{r}
traffic2 <- traffic

```

Now I remove district from the Subagency variable.

```{r}
traffic2$SubAgency <-gsub(" district","",traffic2$SubAgency)
```

Let's see if this worked.

```{r}
head(traffic2)
```

Now that the word district was removed, I'll separate SubAgency into two columns - one containing the district number and the other containing the city.  First I want to see how many SubAgencies there are.

```{r}
table(traffic2$SubAgency)
```

Now let's change the names of these subagencies.  First I check in the crime dataset to make sure we have the same names.

```{r}
#crime <- read_csv("DATASETS\\Crime.csv")
```


```{r}
#table(crime$`Police District Number`)
```
   1D    2D    3D    4D    5D    6D OTHER  TPPD 
10448  9974 15915 14788  9905 12413    18  1771 

I think we will go with the crime dataset names.  I set the Headquarters and Special Operations category to other.

```{r}

traffic2$SubAgency_n[traffic2$SubAgency== "1st, Rockville"] <- "1D"
traffic2$SubAgency_n[traffic2$SubAgency== "2nd, Bethesda"] <- "2D"
traffic2$SubAgency_n[traffic2$SubAgency== "3rd, Silver Spring"] <- "3D"
traffic2$SubAgency_n[traffic2$SubAgency== "4th, Wheaton"] <- "4D"
traffic2$SubAgency_n[traffic2$SubAgency== "5th, Germantown"] <- "5D"
traffic2$SubAgency_n[traffic2$SubAgency== "6th, Gaithersburg / Montgomery Village"] <- "6D"
traffic2$SubAgency_n[traffic2$SubAgency== "Headquarters and Special Operations"] <- "Other"
```

Let's see if this did the trick.

```{r}
head(traffic2)
```

Let's look at the categories

```{r}
table(traffic2$SubAgency)
table(traffic2$SubAgency_n)
```

Looks good.  Now let's see if the class for Date of Stop is correct.

```{r}
class(traffic2$`Date Of Stop`)
```

We need to make sure R knows this is a date.

```{r}
traffic2$Date <- as.Date(traffic2$`Date Of Stop`, "%m/%d/%Y")
```

Let's see if it worked

```{r}
class(traffic2$Date)
```

What about the time of stop?

```{r}
class(traffic2$`Time Of Stop`)
```

R does understand that.  Now let's figure the location variable.  First I create separate columns for year month and day

I may need lubridate so I load it.

```{r}
library(lubridate)
```

Now I create a separate variable for month, day, and year

```{r}
traffic3 <- traffic2
```


```{r}
#Add additional columns for the year, month, and day of the violation
traffic2$year <- as.numeric(format(traffic2$Date,'%Y'))
traffic2$month <- as.numeric(format(traffic2$Date,'%m'))
traffic2$day <- as.numeric(format(traffic2$Date,'%d'))

```


```{r}
head(traffic2)
```

I add the underscores to the columns of the traffic data.

```{r}
colnames(traffic3) = gsub(" ", "_",colnames(traffic3))
```

```{r}
head(traffic3)
```


Ok, Now to the visualization.

```{r}
#these are the furtest latitude/longitude for Montgomery County in any given direction, based on Google Maps
north <- 39.36
south <- 38.92
west <- -77.55
east <- -76.88
ggplot(data = traffic2%>%filter(Longitude>west,Longitude<east,Latitude<north,Latitude>south)) + geom_point(aes(x = Longitude, y = Latitude),alpha = .2, size = .1)+ggtitle("Violations Map")

```

Back to traffic2.  Adding the underscores to the traffic column names.

```{r}
colnames(traffic2) = gsub(" ", "_",colnames(traffic2))
```

Let's see what traffic2 looks like.

```{r}
head(traffic2)
```

Finally, let's see if the Make and Color variables are clean.

```{r}
table(traffic2$Make)
```

The make variable is horible.  Won't be able to use it.  Now let's see color.

```{r}
table(traffic2$Color)
```

This one is more manegeable.

```{r}
traffic2$Color_New <- traffic2$Color
traffic2$Color_New[traffic2$Color_New== "BLUE, DARK"] <- "BLUE"
traffic2$Color_New[traffic2$Color_New== "BLUE, LIGHT"] <- "BLUE"
traffic2$Color_New[traffic2$Color_New== "GREEN, DK"] <- "GREEN"
traffic2$Color_New[traffic2$Color_New== "GREEN, LGT"] <- "GREEN"
traffic2$Color_New[traffic2$Color_New== "GREEN, LGT"] <- "GREEN"
traffic2$Color_New[traffic2$Color_New== "N/A"] <- "OTHER"

```

The renaming seems to have worked as shwon below.

```{r}
table(traffic2$Color_New)
```

```{r}
head(traffic2)
```

Adding a constant column to identify the dataset in case of a merge.

```{r}
traffic2['DATNUM']=1
```

Below I write the dataset to a csv format.

```{r}
head(traffic2)
write.csv(traffic2, "traffic2.csv")
```

traffic3 is the same as traffic 2 except that it includes a constant to identify the traffic violations as 1.

```{r}
traffic3['DATNUM']=1
traffic3['V_type']=1
traffic4 <- traffic3

```

traffic4 is a subset of the traffic3 with only select variables.

```{r}
traffic4 <- traffic4 %>%
  select(Date, DATNUM, Latitude, Longitude, SubAgency_n, V_type)
```



```{r}
table(traffic2$SubAgency_n)
```

```{r}
table(traffic2$SubAgency)
```


```{r}
table(traffic2$Race)
```


```{r}
traffic2$RACE <- traffic2$Race
traffic2$RACE[traffic2$Race== "NATIVE AMERICAN"] <- "OTHER"
table(traffic2$RACE)
```


```{r}
head(traffic2)
```

```{r}
# Load function
source("http://pcwww.liv.ac.uk/~william/R/crosstab.r")
```


```{r}
table(traffic2$Contributed_To_Accident)
```

```{r}
library(ggthemes)
```


```{r}
traffic2$Gender_n <- traffic2$Gender
traffic2$Gender_n[traffic2$Gender_n == "U"] <- NA
table(traffic2$Gender_n)
```

```{r}
trafficS <- traffic2
```


```{r}
traffic2$RACE_n <- factor(traffic2$RACE, levels = c("WHITE", "BLACK", "HISPANIC", "ASIAN", "OTHER"))
table(traffic2$RACE_n)
table(traffic2$RACE)
```


```{r}
options(scipen=10000)
traffic2 %>%  filter(!is.na(Gender_n)) %>%
ggplot() +
aes(x = RACE_n , fill = Gender_n) +
geom_bar(position = 'dodge') +
scale_fill_discrete(labels = c("Female", "Male")) +
labs(fill = "Gender", y = "# of stops", x ="Race", title = "Number of Stops by Race and Gender") +
theme_economist_white()

```

MD estimates are 43 % white, 20% black, 20% hispanic, 16% Asian

```{r}
43+20+20+16
```


```{r}
traffic2 %>%  filter(Alcohol == "Yes") %>%  filter(!is.na(Gender_n)) %>%
ggplot() +
aes(x =  RACE_n, fill = Gender_n) +
geom_bar(position = 'dodge')  +
scale_fill_discrete(labels = c("Female", "Male")) +
labs(fill = "Gender", y = "# of stops", x ="Race", title = "Number of Stops Involving Alcohol by Race and Gender") +
theme_economist_white()
```


```{r}
traffic2 %>%  filter(Contributed_To_Accident == "Yes") %>%  filter(!is.na(Gender_n)) %>%
ggplot() +
aes(x =  RACE, fill = Gender_n) +
geom_bar(position = 'dodge')  +
scale_fill_discrete(labels = c("Female", "Male")) +
labs(fill = "Gender", y = "# of stops", x ="Race", title = "Accidents by Race and Gender") +
theme_economist_white()
```


```{r}
traffic2 %>%  filter(Contributed_To_Accident == "Yes") %>%  filter(!is.na(Gender_n)) %>% filter(Alcohol == "Yes") %>%
ggplot() +
aes(x =  RACE) +
geom_bar(position = 'dodge')  +
labs(fill = "Gender", y = "# of stops", x ="Race", title = "Accidents Involving Alcohol") +
theme_economist_white()
```

```{r}
table(traffic2$SubAgency)
```



```{r}
traffic2  %>% filter(!is.na(SubAgency_n)) %>% filter(SubAgency_n!= "Other") %>%
ggplot() +
aes(x =  SubAgency_n) +
geom_bar(position = 'dodge')  +
labs(y = "# of stops", x ="District", title = "Number of Stops by District") +
theme_economist()
```



```{r}
traffic2 %>%  filter(Alcohol == "Yes") %>%  filter(SubAgency_n!= "Other") %>%
ggplot() +
aes(x =  SubAgency_n) +
geom_bar(position = 'dodge')  +
labs(y = "# of stops", x ="District", title = "Number of Stops Involving Alcohol by District") +
theme_economist()
```


```{r}
table(traffic2$Color_New)
```

```{r}

traffic2$Color_n <- traffic2$Color_New
traffic2$Color_n[traffic2$Color_n == "GRAY"] <- "SILVER"
traffic2$Color_n[traffic2$Color_n != "GRAY" & traffic2$Color_n != "BLACK" & traffic2$Color_n != "WHITE" & traffic2$Color_n != "BLUE" & traffic2$Color_n != "SILVER" & traffic2$Color_n != "RED"] <- "OTHER"

```

```{r}
table(traffic2$Color_n)
```



```{r}
traffic2 %>% filter(RACE!="OTHER") %>%
ggplot()+
aes(x=Time_Of_Stop, colour=RACE) + geom_bar() + labs(colour= "Race", y = "# of stops", x ="Time", title = "Number of Stops by Time")  + scale_colour_discrete(labels = c("Asian", "Black", "Hispanic", "White")) +
theme_stata()

```



```{r}
#traffic2 %>% 
#filter(traffic2$RACE != "OTHER") %>%

traffic2 %>% filter(RACE != "OTHER") %>% filter(SubAgency_n != "Other") %>% crosstab(row.vars = "RACE", col.vars = "SubAgency_n", type = "f")
traffic2 %>% filter(RACE != "OTHER") %>% filter(SubAgency_n != "Other") %>% crosstab(row.vars = "RACE", col.vars = "SubAgency_n", type = "c")
traffic2 %>% filter(RACE != "OTHER") %>% filter(SubAgency_n != "Other") %>% crosstab(row.vars = "RACE", col.vars = "SubAgency_n", type = "r")
#crosstab(traffic2, row.vars = "bias_cat", col.vars = "Bias_n", type = "c")
#crosstab(traffic2, row.vars = "bias_cat", col.vars = "Bias_n", type = "r")
```

```{r}
table(traffic2$SubAgency)
```


```{r}
traffic2 %>%  filter(RACE != "OTHER") %>% filter(SubAgency_n != "Other") %>%
ggplot() +
aes(x =  SubAgency_n, fill = RACE) +
geom_bar(position = 'dodge')  +
scale_fill_discrete(labels = c("Asian", "Black", "Hispanic", "White")) +
labs(fill = "Race", y = "# of stops", x ="District", title = "Accidents by Race and Gender") +
theme_economist_white()
```

```{r}
traffic2 %>%  filter(RACE != "OTHER") %>% filter(SubAgency_n != "Other") %>%
ggplot() +
aes(x =  RACE, fill = SubAgency_n) +
geom_bar(position = 'dodge') +
labs(fill = "District", y = "# of stops", x ="Race", title = "Accidents by Race and Gender") +
theme_economist_white()
```


```{r}
table(traffic2$Fatal)
```


```{r}

crosstab(traffic2, row.vars = "Contributed_To_Accident", col.vars = "Alcohol", type = "f")
crosstab(traffic2, row.vars = "Contributed_To_Accident", col.vars = "Alcohol", type = "c")
crosstab(traffic2, row.vars = "Contributed_To_Accident", col.vars = "Alcohol", type = "r")
```


```{r}
table(traffic2$RACE)
```

```{r}
table(traffic2$Gender)

```

```{r}
table(traffic2$Gender_n)
```


```{r}
table(traffic2$SubAgency)
table(traffic2$SubAgency_n)
```

```{r}
table(traffic2$Color_n)
```



